Quantitative Auditing of Fairness Measures with Differentially Private Synthetic Data

### 2024-12-02 paper/paper.tex
some progress

Fairness in machine learning has become a crucial area of research\cite{barocas2023fairness}. Various fairness measures have been proposed to quantify the fairness of AI models\cite{pessach2022review,corbett2017algorithmic,vzliobaite2017measuring,hardt2016equality,corbett2017algorithmic,berk2021fairness,chouldechova2017fair,kleinberg2016inherent}.

@@

The generation of synthetic data under differential privacy constraints allows for data sharing without compromising privacy\cite{tao2021benchmarking}. There have been several techniques developed for Differentially Private synthetic data generation\cite{rosenblatt2020differentially,fan2020relational,bowen2019comparative,bowen2020comparative,arnold2020really,xu2019modeling} in recent years.

While useful, there are some ethical concerns with the generation of synthetic data ...

% TODO: cite

@@

Let $\prod_i x_i$ denotes the Cartesian product of $x_i$s.

@@

The task of differentially private synthetic data\cite{Ullman2022,McKenna2022,NIST_Differentially_Private_Synthetic_Data,YouTube_Differentially_Private_Synthetic_Data} is, given a database $\db$, adding some noise to marginals of $\db$ such that it satisfies some Differential Privacy guarantees and outputting another database $\db'$, such that the $L^1$-norm errors between marginals of $\db$ and $\db'$ is small; that is, their marginals $\mu_{\db},\mu_{\db'}$ are similar.

>>>>

Fairness in machine learning has become a crucial area of research\cite{barocas2023fairness}. Various fairness measures have been proposed to quantify\cite{yeh2024analyzing} the fairness of AI models\cite{pessach2022review,corbett2017algorithmic,vzliobaite2017measuring,hardt2016equality,corbett2017algorithmic,berk2021fairness,chouldechova2017fair,kleinberg2016inherent}.

@@

The generation of synthetic data\cite{raghunathan2021synthetic,lu2023machine} under Differential Privacy constraints allows for data sharing without compromising privacy\cite{tao2021benchmarking}. There have been several techniques developed for Differentially Private synthetic data generation\cite{rosenblatt2020differentially,fan2020relational,bowen2019comparative,bowen2020comparative,arnold2020really,xu2019modeling} in recent years.

There are also works that aim to introduce bias to synthetic data\cite{jiang2024synthetic,baumann2023bias}.

While helpful, there are some technical limitations\cite{stadler2022synthetic,cheng2021can,ganev2022robin,wyllie2024fairness} and ethical concerns\cite{whitney2024real} with the use of synthetic data.

@@

Let $\prod_i x_i$ denote the Cartesian product of $x_i$s.

@@

The task of Differentially Private synthetic data\cite{Ullman2022,McKenna2022,NIST_Differentially_Private_Synthetic_Data,YouTube_Differentially_Private_Synthetic_Data} is, given a database $\db$, adding some noise to marginals of $\db$ such that it satisfies some Differential Privacy guarantees and outputting another database $\db'$, such that the $L^1$-norm errors between marginals of $\db$ and $\db'$ is small; that is, their marginals $\mu_{\db},\mu_{\db'}$ are similar.
